---
title: "International Cherry Blossom Prediction Competition"
subtitle: "Data 2024"
author: "Edimer David Jaramillo"
date: "February 29, 2024"
lang: en-US
format:
  html:
    page-layout: article
    toc: true
    code-fold: true
    df-print: paged
    toc-location: left
    number-depth: 4
    theme: yeti
    code-copy: true
    highlight-style: github
    embed-resources: true
    code-tools:
      source: true    
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      error = FALSE, 
                      message = FALSE,
                      fig.align = 'center')
```

# Libraries and setup

```{r}
# Libraries
library(tidyverse)
library(arrow)
library(glue)
library(furrr)
library(splines)

# Colors
colors_custom <-
  c("#014e25",
    "#800080",
    "#ffa500",
    "#008080",
    "#ff6347",
    "#0000cd")

# Theme ggplot2
theme_set(theme_bw() + theme(legend.position = "top"))

# Functions
fs::dir_ls("../source/r/module-models/") |>
  walk(.f = source) 
  

# Inputs
countries <- c("Japan", "Switzerland", "South Korea", "USA-WDC")
```

# Initial data

- Filter information older than the year 1980 for all databases.
- I do not take into account the `shrubs` variable due to lack of data.
- To avoid problems with the coordinates, I round the latitude and longitude to 4 digits.
- I create a variable named `id_coord` to identify each coordinate and facilitate some calculations.
- In this execution I only take into account the coordinates of interest to predict. In order to execute the functions without problem, I must assume a reference date to extract summaries prior to that date, as explained in the document **07-Modeling-FeatEnginer-FeatSelection.qmd**, the date I use in this case is "2024-03-25", I choose this date to guarantee that I will have data from moments before flowering.


```{r}
#| eval: false

# ---- Initial data ----
data_coords_predict <-
  read_parquet("../external-data/data-models/df_coords_predict.parquet") |>
  mutate(lat = round(lat, digits = 4),
         long = round(long, digits = 4))

df_bloom_complete <-
  read_parquet("../external-data/data-models/df_full_complete.parquet") |>
  filter(year > 1980) |>
  mutate(
    lat = round(lat, digits = 4),
    long = round(long, digits = 4),
    id_coord = str_c(lat, "-", long)
  ) |>
  filter(lat %in% data_coords_predict$lat) |>
  distinct(lat, long, .keep_all = TRUE) |>
  mutate(bloom_date = ymd("2024-03-25"))

df_weather <-
  read_parquet("../external-data/df_weather.parquet") |>
  mutate(
    lat = round(lat, digits = 4),
    long = round(long, digits = 4),
    id_coord = str_c(lat, "-", long)
  )

df_photoperiod <-
  read_parquet("../external-data/df_photoperiod.parquet") |>
  filter(year(date_photoperiod) > 1980) |>
  mutate(lat = round(lat, digits = 4))

# Feature Engineering

# ---- Predictors ----

## Fixed ----
data_predictors_fixed <-
  df_bloom_complete |>
  distinct(lat, long, .keep_all = TRUE) |>
  select(-c(location, year, bloom_date, bloom_doy, shrubs, id_coord)) |>
  distinct(lat, long, .keep_all = TRUE) |>
  janitor::clean_names()

write_parquet(
  data_predictors_fixed,
  "../external-data/data-predict/data_predictors_fixed.parquet"
)

## Summary predictors with climate ----
options(future.globals.maxSize = 1000 * 1024 ^ 2) # Config furrr
plan(multisession, workers = parallel::detectCores() - 2)

### 1 Month ----
number_months_weather <- 1

data_predictors_summary_weather1M <-
  FEWeatherSummary(
    months_window = number_months_weather,
    data_bloom = df_bloom_complete,
    data_weather = df_weather
  )

write_parquet(
  data_predictors_summary_weather1M,
  "../external-data/data-predict/data_predictors_summary_weather1M.parquet"
)

### 3 Months ----
number_months_weather <- 3

data_predictors_summary_weather3M <-
  FEWeatherSummary(
    months_window = number_months_weather,
    data_bloom = df_bloom_complete,
    data_weather = df_weather
  )

write_parquet(
  data_predictors_summary_weather3M,
  "../external-data/data-predict/data_predictors_summary_weather3M.parquet"
)

### 6 Months ----
number_months_weather <- 6

data_predictors_summary_weather6M <-
  FEWeatherSummary(
    months_window = number_months_weather,
    data_bloom = df_bloom_complete,
    data_weather = df_weather
  )

write_parquet(
  data_predictors_summary_weather6M,
  "../external-data/data-predict/data_predictors_summary_weather6M.parquet"
)

### 9 Months ----
number_months_weather <- 9

data_predictors_summary_weather9M <-
  FEWeatherSummary(
    months_window = number_months_weather,
    data_bloom = df_bloom_complete,
    data_weather = df_weather
  )

write_parquet(
  data_predictors_summary_weather9M,
  "../external-data/data-predict/data_predictors_summary_weather9M.parquet"
)

### 12 Months ----
number_months_weather <- 12

data_predictors_summary_weather12M <-
  FEWeatherSummary(
    months_window = number_months_weather,
    data_bloom = df_bloom_complete,
    data_weather = df_weather
  )

write_parquet(
  data_predictors_summary_weather12M,
  "../external-data/data-predict/data_predictors_summary_weather12M.parquet"
)

## Summary predictors with photoperiod ----
options(future.globals.maxSize = 1000 * 1024 ^ 2) # Config furrr
plan(multisession, workers = parallel::detectCores() - 2)

### 1 Month ----
number_months_photo <- 1

data_predictors_summary_photo1M <-
  FEPhotopSummary(months_window = number_months_photo,
                  data_bloom = df_bloom_complete,
                  data_photo = df_photoperiod)

write_parquet(
  data_predictors_summary_photo1M,
  "../external-data/data-predict/data_predictors_summary_photo1M.parquet"
)

### 3 Months ----
number_months_photo <- 3

data_predictors_summary_photo3M <-
  FEPhotopSummary(months_window = number_months_photo,
                  data_bloom = df_bloom_complete,
                  data_photo = df_photoperiod)

write_parquet(
  data_predictors_summary_photo3M,
  "../external-data/data-predict/data_predictors_summary_photo3M.parquet"
)

### 6 Months ----
number_months_photo <- 6

data_predictors_summary_photo6M <-
  FEPhotopSummary(months_window = number_months_photo,
                  data_bloom = df_bloom_complete,
                  data_photo = df_photoperiod)

write_parquet(
  data_predictors_summary_photo6M,
  "../external-data/data-predict/data_predictors_summary_photo6M.parquet"
)

### 9 Months ----
number_months_photo <- 9

data_predictors_summary_photo9M <-
  FEPhotopSummary(months_window = number_months_photo,
                  data_bloom = df_bloom_complete,
                  data_photo = df_photoperiod)

write_parquet(
  data_predictors_summary_photo9M,
  "../external-data/data-predict/data_predictors_summary_photo9M.parquet"
)

### 12 Months ----
number_months_photo <- 12

data_predictors_summary_photo12M <-
  FEPhotopSummary(months_window = number_months_photo,
                  data_bloom = df_bloom_complete,
                  data_photo = df_photoperiod)

write_parquet(
  data_predictors_summary_photo12M,
  "../external-data/data-predict/data_predictors_summary_photo12M.parquet"
)

## Summary predictors with weather (GDD) ----
options(future.globals.maxSize = 1000 * 1024 ^ 2) # Config furrr
plan(multisession, workers = parallel::detectCores() - 2)

data_predictors_GDD <-
  featureEngGDD(
    start_date = "01-01",
    # MMDD
    t_basal = 5,
    # °C
    data_bloom = df_bloom_complete,
    data_weather = df_weather
  )

write_parquet(
  data_predictors_GDD,
  "../external-data/data-predict/data_predictors_GDD.parquet"
)
```

# Complete data: predict 2024

- I join the databases `data_predictors_fixed`, `data_predictors_summary_weather[w]M`, `data_predictors_summary_photo[w]M` and `data_predictors_gdd`. This is the final database to make predictions for the year 2024.

```{r}
#| eval: false

# Auxiliar function ----
extract_number <- function(text) {
  match <- gregexpr("\\d+", text)
  number <- regmatches(text, match)
  number <- as.integer(number)
  return(number)
}

# =================== PREDICTORS =========================================

# Import data: 36 predictors ----
data_predictors_fixed <- 
  read_parquet("../external-data/data-predict/data_predictors_fixed.parquet")

## Weather: 180 predictors ----
data_predictors_summary_weather <-
  fs::dir_ls("../external-data/data-predict/", regexp = "summary_weather") |>
  map(
    .f = function(x = .x) {
      read_parquet(x) |>
        mutate(month_weather = str_c("month_", extract_number(x)))
    }
  ) |>
  list_rbind() |>
  pivot_wider(
    names_from = month_weather,
    values_from = -c(lat, long, month_weather, year_bloom)
  ) |>
  unnest() |>
  distinct_all()

## Photoperiod: 15 predictors ----
data_predictors_summary_photo <-
  fs::dir_ls("../external-data/data-predict/", regexp = "summary_photo") |>
  map(
    .f = function(x = .x) {
      read_parquet(x) |>
        mutate(month_photo = str_c("month_", extract_number(x)))
    }
  ) |>
  list_rbind() |>
  pivot_wider(
    names_from = month_photo,
    values_from = -c(lat, month_photo, year_bloom)
  ) |>
  unnest() |>
  distinct_all()

## AGDD: 3 predictors ----
data_predictors_gdd <-
  read_parquet("../external-data/data-predict/data_predictors_GDD.parquet")

# =================== TARGET =========================================
data_modeling <-
  df_bloom_complete |>
  select(location, lat, long, year_bloom = year, bloom_doy) |>
  left_join(data_predictors_fixed, by = c("lat", "long")) |>
  left_join(data_predictors_summary_weather,
            by = c("lat", "long", "year_bloom")) |>
  left_join(data_predictors_summary_photo, by = c("lat", "year_bloom")) |>
  left_join(data_predictors_gdd, by = c("lat", "long", "year_bloom")) |>
  relocate(location, country, everything()) |>
  distinct(lat, long, year_bloom, .keep_all = TRUE)

write_parquet(data_modeling,
              "../external-data/data-predict/data_predictors_2024.parquet")
```
